---
title: "Prototype - An agent-based platform to evaluate the performance of sampling designs and modelling approaches when estimating animal abundance"
author: "Denner A. B. Fabiano, Iuri E. P. Ferreira"
date: "27/10/2021"
output: 
  html_document: 
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r}
#install.packages("landscapeR")
#install.packages("raster")
library(raster) 
library(landscapeR)
```

## To generate the landscape

Função que cria o ambiente (paisagem). O usuário pode configurar o tamanho da rede ``l.size``, o *vetor* número de fragmentos ``n.patches``, com três níveis de recurso (pobre, intermediário, rico), e cobertura de cada tipo de fragmento ``cover`` (soma da ``cover`` < 1).   

 + default para ``l.size`` é 100 px,
 + default para ``n.patches`` é 2 fragmentos de cada tipo,
 + default para ``cover`` é de 25% para cada fragmento.   


```{r}
land.create <- function(l.size = 100, n.patches = c(2, 2, 2), cover = c(0.25, 0.25, 0.25)){
  
  # Cria valores de qualidade / recursos 
  val <- c(1, 3, 7)
  
  # Cover = cobertura de cada tipo de paisagem (soma < 1)
  if( sum(cover) > 1 ){print("Erro. Cobertura total não pode ser maior do que 1"); break}
  
  # Total de recursos 
  t.r <- val %*% cover
  
  #### Criando a matriz de recurso ####
  
  # criando uma paisagem vazia - só com zeros
  m <- matrix(0, l.size, l.size) # the empty landscape
  
  # transformando a matriz em raster
  r <- raster(m, xmn=0, xmx=l.size, ymn=0, ymx=l.size)
  
  # cria sobre o raster r manchas disjuntas que 
  # representam fragmentos pobres (i=1), 
  # intermediários (i=2) e ricos (i=3)
  for(i in 1:length(cover)){
    r <- makeClass(r, n.patches[i],
                   round(cover[i]*l.size^2/n.patches[i]), val = val[i])
  }
  
  # retorna o que é importante
  return( list("raster"=r,"recursos"=t.r) )
  
} # fim da função land.create
# teste
obj <- land.create(200, c(1, 2, 3), c(0.20,0.15,0.30))
# exemplo: criar uma paisagem grande, pobre e muito fragmentada
obj <- land.create(l.size = 500, n.patches = c(10, 3, 3), 
                   cover = c(0.30, 0.05, 0.05) )
plot(obj$raster)
```


## Solitary animals / population sim. 

Each time step is one hour. Simulations will represent 21 consecutive weeks. We considered the first twenty weeks needed to individuals establish the natural pattern of occupancy from random initial positions (adaptation phase). The remaining week is used to simulate the deposition of traces on the environment and level of patch occupancy. 

We considered the individual rate of four traces produced every day.   

 + ``n`` is the number of animal on the environment
 + ``ra`` é o argumento que recebe o raster da paisagem
 + ``time`` is the number of weeks (adaptation + study phases)
 + ``where`` is where the animals are placed at the start of simulation (over all the environment or only where there are resources)

```{r}
animal.mov<-function( n = 100, ra = NULL, time = 21, where = "resource"){

  # Time in weeks (20 adaptation + 1 experiment)
if(time<=1){print("Error. At least two weeks are necessary!"); break}
  
time.step<-round(time)*7*24
time.step.adapt<-(round(time)-1)*7*24

# Número de animais (n)
# r = valores do raster (formato de matriz)
if(is.null(ra)){print("Erro. Inserir os valores do raster no segundo argumento"); break}
  
r<-values(ra,format="matrix")

l<-dim(r)[1]

if(where=="random"||where=="resource"){
if(where=="random"){
M<-matrix(sample(1:l,2*n,replace=TRUE),ncol=2)
} else{
coords<-NULL
for(i in 2:(l-1)){
for(j in 2:(l-1)){
if(r[i,j]>0){coords<-rbind(coords,c(i,j))}  
}  
}
size<-dim(coords)[1]  
k<-sample.int(size,n,replace=TRUE)  
M<-coords[k,]  # matriz de posição dos animais
} # fim else
} else{
print("Error. where must be random or resource"); break  
}  
  
R<-NULL
MAUX<-M
R[[1]]<-M

classe<-c(1,2,3,4,5,6,7,8,9)
traces<-NULL
for(j in 2:time.step){ 
# Atualização 
for(i in 1:n){ 
x<-M[i,1]; y<-M[i,2]

if(j>time.step.adapt&j%%6==0){traces<-rbind(traces,c(x,y))}

if(x==1|x==l|y==1|y==l){
yaux<-y; xaux<-x
  
  if(x==1){xaux<-x+1}
  if(x==l){xaux<-x-1}
  
  if(y==1){yaux<-y+1}
  if(y==l){yaux<-y-1}
 
MAUX[i,]<-c(xaux,yaux)
} else{
# atualização dentro da matriz
v.aux<-c(r[x-1,y+1],r[x,y+1],r[x+1,y+1], 
r[x-1,y],r[x,y],r[x+1,y],
r[x-1,y-1],r[x,y-1],r[x+1,y-1])
if(sum(v.aux)>0){
p.aux<-v.aux/sum(v.aux)
cl<-sample(classe,1,prob=p.aux)
} else{
  cl<-sample(classe,1)
} 
  
switch(cl,
       MAUX[i,]<-c(x-1,y+1), 
       MAUX[i,]<-c(x,y+1),
       MAUX[i,]<-c(x+1,y+1),
       MAUX[i,]<-c(x-1,y),
       MAUX[i,]<-c(x,y),
       MAUX[i,]<-c(x+1,y),
       MAUX[i,]<-c(x-1,y-1),
       MAUX[i,]<-c(x,y-1),
       MAUX[i,]<-c(x+1,y-1),
       )

} # fim else 
} # fim i
M<-MAUX      
R[[j]]<-M
} # fim j 
return(list("Evol"=R,"Traces"=traces,"time.step"=time.step,"time.step.adapt"=time.step.adapt))
}
```
 
## Individual tracks' viewer 

The track is registered hour by hour. 
 
```{r}
# To insert the output of the animal.mov function
tracks.viewer<-function(obj=NULL, r=NULL, adapt=FALSE){
if(is.null(obj)){print("Erro. Inserir os resultados da simulação"); break}
# r = valores do raster (formato de matriz)
if(is.null(r)){print("Erro. Inserir os valores do raster no segundo argumento"); break}

# number of animals / tracks
n<-dim(obj$Evol[[1]])[1]
time.step<-obj$time.step
time.step.adapt<-obj$time.step.adapt
tracks<-NULL
for(i in 1:n){
coords<-NULL
for(j in 1:time.step){
coords<-rbind(coords,obj$Evol[[j]][i,])
}
tracks[[i]]<-coords
}
if(adapt==FALSE){
plot(flip(t(r),direction='y'))+for(i in 1:n) lines(tracks[[i]][(time.step.adapt+1):time.step,],type='l',cex=0.05,col=1)+points(tracks[[i]][time.step,1],tracks[[i]][time.step,2],cex=0.8,col=2,
        pch=19)} else{
plot(flip(t(r),direction='y'))+for(i in 1:n) lines(tracks[[i]][1:time.step.adapt,],type='l',cex=0.05,col=4) + lines(tracks[[i]][(time.step.adapt+1):time.step,],type='l',cex=0.05,col=1)+
points(tracks[[i]][time.step,1],tracks[[i]][time.step,2],cex=0.8,col=2,pch=19)  
}
} # fim function
```

## Viewer for traces

```{r}
# To insert the output of the animal.mov function
traces.viewer<-function(obj=NULL, r=NULL){
if(is.null(obj)){print("Erro. Inserir os resultados da simulação"); break}
# r = valores do raster (formato de matriz)
if(is.null(r)){print("Erro. Inserir os valores do raster no segundo argumento"); break}
#flip(t(rtest),direction='y')
plot(flip(t(r),direction='y'))+points(obj$Traces,cex=0.3,pch=19,col=1)
}  
```
## Make Line transects 
```{r}
tracks.set<-function(obj=NULL,ra=NULL,l=30,sampling="random",dir=NULL,loc=NULL, con=0.01){
# ntracks = number of transects
# l = length of transects 
# w = width of transects  
# r = valores do raster (formato de matriz)
if(is.null(obj)){print("Erro. Inserir o objeto tipo animal.mov"); break}

  
if(is.null(ra)){print("Erro. Inserir os valores do raster no segundo argumento"); break}

ll<-dim(ra)[1]
    
if(is.null(dir)){dir<-runif(1,0,360)}
if(is.null(loc)){loc<-round(runif(2,ll/9,ll*8/9))}  

if(sampling=="random"||sampling=="convenience"){
if(sampling=="random"){
# spt = NULL -> then random locations
m <- matrix(0, ll, ll) # the empty landscape
r <- raster(m, xmn=0, xmx=ll, ymn=0, ymx=ll)
rtest<-makeLine(r, l, direction = dir, convol = con, spt = loc, edge = FALSE, bgr=0, rast = TRUE, val = 10)
} else{
rtest<-makeLine(ra, l, direction = dir, convol = con, spt = loc, edge = FALSE, bgr=7, rast = TRUE, val = 10)  
} # end first else
} else{
  print("Error. Sampling design must be at random or by convenience")
} # end second else


area.t<-sum(values(rtest,format="matrix")==10)
coverage<-area.t/(ll^2)
coord<-NULL
for(i in 1:ll){
for(j in 1:ll){
if(rtest[i,j]==10){coord<-rbind(coord,c(i,j))}
}
}

traces<-din$Traces
counts<-0
for(k in 1:dim(traces)[1]){
if(rtest[traces[k,1],traces[k,2]]==10){counts<-counts+1}  
}

plot(flip(t(ra),direction='y'))+points(coord[,1],coord[,2],col="blue",pch=15,type="p",xlim=c(0,ll),ylim=c(0,ll),cex=0.5)+points(traces,pch=15,cex=0.2)
return(list("transect"=rtest,"coverage"=coverage,"area"=area.t,
            "counts"=counts))
}
```

## Basic Workflow

### To create scenarios / landscapes
```{r}
# l.size de 500 = 50 segundos de simulação
land<-land.create(l.size=500,n.patches=c(3,3,3),cover=c(0.15,0.20,0.25))
plot(land$raster)
```

### To simulate moviments 

```{r}
rm(list="ra")
ra<-land$raster
(pop.size<-round(land$recursos*20)) # 20 foi arbitrário
# Execução da rotina
start_time <- Sys.time()
din<-animal.mov(n=pop.size,ra=ra,time=21,where="resource")
end_time <- Sys.time()
end_time - start_time
# traces positions
#din$Traces
```

### To visualize tracks
```{r}
tracks.viewer(obj=din,r=ra,adapt=FALSE)
```

### To visualize traces in environment

```{r}
traces.viewer(obj=din,r=ra)
```

### To set sampling transects and to count traces
```{r}
tracks<-tracks.set(obj=din,ra=ra,sampling="random",
                   l=300,con=0.01)
tracks$coverage
tracks$counts
```









